---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# cwl - Common workflow language support

This package (cwl) is designed to parse and work with the Common Workflow Language, CWL. ( http://www.commonwl.org). It currently does very little, it is only a stub in which various representations and algorithms are being played with. The first 0.0.1 release will be able to run simple command line programs locally using a very limited subset of CWL.

## Simple command line tool examples.

These assume you already have a _tool.cwl_ and a _data.yaml_ file. See the introductory vignette for details and sample files.

```{r, eval=FALSE}
# Run a tool directly from files
cwlRun( "tool.cwl", "data.yaml")

# Run a tool indirectly via object intermediaries.
# Allows interaction with the tool and data info before running.
toolObj <- readCwlTool( "tool.cwl" )
dataObj <- readCwlData( "data.yaml" )
cwlRun( toolObj, dataObj)
```

